\chapter{About the author}
\epigraph{Hakuna Matata}{Timon and Pumbaa\\\textit{The Lion King}}

Bram was born in 1987 Ramotswa, Botswana. Fast forwarding a bit, he studied Electrical Engineering at the University of Twente, the Netherlands. During his studies he was an active member of the study association of Electrical Engineering, E.T.S.V.~Scintilla. Part of his activities included: being board member, participating in the study tour, and partaking in various committees. His graduation project (2012) focused on exploring the usability of finger vein biometrics~\footnote{\fullcite{ton13}}. A sensor was designed to capture finger vein images and several feature extraction and matching algorithms were evaluated.

Fast forwarding again, in 2015 he was part of a start-up, BallJames which was a subsidiary of SciSports. Instead of digitising rail environments, this start-up aimed at digitising football matches using several synchronised cameras around the pitch. Using computer vision techniques and deep learning models, the ultimate goal would be to know exactly when and where each player was during the match.

In 2018 he started working at the Saxion University of Applied Sciences. He started his career at the Electrical Engineering and Applied Computer Sciences department. He was involved in teaching courses such as Digital Signal Processing and the basics of digital techniques. After some time he joined the Ambient Intelligence research group at Saxion, and gradually moved more towards research. During this time also an opportunity opened up to pursue a PhD in joint collaboration with the University of Twente which he decided to take. And the rest is history so to say.

In his free time he enjoys tinkering with his FreeBSD home server, attempting some wood working, and creating the occasional pull request for an open source project.