\chapter{Acknowledgements}
\epigraph{It is never finished, but there is an end.}{Sam Aerts\\\textit{The Hague University of Applied Sciences}}

In my opinion, the above epigraph is an adequate summary of doing a PhD. During this wild roller-coaster ride, the numerous interactions with various people have helped me gain new insights and ideas which contributed to the completion of this thesis. So, without further ado, my feeble attempt at giving due credit:

First of all, my employer, Saxion the University of Applied Sciences. And specifically the Ambient Intelligence research group for providing me with the opportunity, freedom, and resources to pursue this unique journey. A very warm thank you to all my wonderful colleagues from our Ambient Intelligence research group at Saxion, you are really the ones which make going to work a pleasure. Furthermore, the large number of students, both from Saxion and the University of Twente which I have supervised. The supervision process and the challenging questions from alternative viewpoints have always been a big bonus for my personal knowledge gain. In addition, I would like to thank my promotor Paul and my daily supervisor Berend~Jan from the Pervasive Systems group at the University of Twente, for the talks and discussions.

During my research I have also worked closely together with Strukton Rail. Especially I would like to thank the following people from Strukton Rail for the nice and easy-going collaboration: Corn\'{e}, Renzo, Rick, Jesse, and Wouter. Thank you (?) for revealing the underlying complexity of maintaining and managing such a complex and vast infrastructure. With the knowledge and insights gained from our interactions, travelling by train will never be the same any more.

In addition, I would like to thank some bands\footnotemark, in my opinion they certainly deserve their fair share of appraisal. Their music, combined with copious amounts of black coffee, have helped me stay focused during coding and writing sessions, and resulted into this nice looking thesis right in front of you.

I would also like to use this opportunity to thank two other social groups which have been very dear to me. First are the `stam' members of scouting Euregio, although I don't see you as often any more, I still cherish the memories of our time together during large part of my adolescence. Secondly, the other members of EHBD, which are Erwin, Harm and Denick. Having peers to share my IP networking woes with is very comforting, I can tell you.

Furthermore, I would like to thank Lesley. I could not find a fitting category for you, so here it is. It is always nice to meet you and I enjoy our discussions on a wide variety of topics. I wish you all the best with your upcoming fatherhood, as gamers (used to ?) say: GLHF!

Impossible to forget: my parents, for an incredible and unique childhood, and their continued unconditional support. And a special thanks to Tias, whom I have known since `middelbare school', for discussing all matters of life and watching/critiquing recent TV series together.

And last, but not least, my wife, and my lovely two children, Willem and Mette. At the end of the day, you are always the ones able to `turn my frown upside down'.

% Or \raggedleft
\vspace{2cm}
\noindent Bram\\
Hengelo, November 2024

\footnotetext{In no particular order, besides alphabetically: Amaranthe, BABYMETAL, Bad Religion, Blink-182, Disturbed, Five Finger Death Punch, Greenday, Heideroosjes, Linkin Park, Metallica, Offspring, Papa Roach, Rammstein, Red Hot Chili Peppers, Slipknot, and Soulfly.}