\chapter{Immersive visualisation}\label{chap:immvis}
\epigraph{Visualisations act as a campfire around which we gather to tell stories.}{Al Shalloway\\\textit{Founder of Net Objectives}}
\blfootnote{This chapter has been previously published as:\\ \fullcite{ton2024imvis}}

The previous two chapters have focused on extracting information from point cloud data, thus enabling the transition to cyber space. To complete the cycle, by going back to the physical world, this chapter will focus on how this information can be visualised in an immersive environment. Benefits of immersive visualisation are the reduced time to insight, the intuitive user experience, and the possibility of collaborative viewing.

Despite these benefits, immersive visualisation of point cloud data for the rail domain has not been explored yet. To bridge this gap, this chapter presents a working prototype to immersively visualise railway point cloud data. The prototype enables the user to load a point cloud scene and interact with the scene using natural hand gestures and voice commands. Manual interactions include scaling, rotating and translating the scene. This prototype is one of the contributions of this thesis.

Based on a small user evaluation, it was found that participants had difficulty with the voice commands to control the visibility of objects. Key focus points for a next iteration are dynamically loading large scenes with an adaptive level of detail and expanding hand gesture based interaction. The work presented in this chapter is used to answer RQ3.

\clearpage
\input{./Chapters/immvis/immvis.tex}

\clearpage
\printbibliography[heading=subbibnumbered]