\chapter{State of the art}\label{chap:stoa}
\epigraph{The more I see, the less I know.}{Snow (Hey Oh)\\\textit{Red Hot Chili Peppers}}
\blfootnote{This chapter has been previously published as:\\ \AtNextCite{\defcounter{minnames}{3}}\fullcite{dekker23}}

This chapter serves as a snapshot of the current state of the art addressing the pivotal role of point cloud techniques in advancing railway digitalisation and providing valuable pointers for future research directions.

Employing a systematic review approach, this chapter concentrates exclusively on research centred around railway assets and their digitalisation via point cloud data. The literature is themed into pre-processing, modelling, and digital twinning. 

This chapter analyses diverse modelling and pre-processing techniques and categorises them for clarity. Digital twinning techniques are also collected, though these techniques are scarce in the context of railway infrastructure and point clouds. This chapter also presents a compilation of dataset statistics highlighting the scarcity of openly available railway-specific datasets. This scarcity considerably hampers the feasibility of research reproducibility and the comparative analysis of different approaches.

Particularly, it is conclude that hybrid methodologies that combine machine learning with structure-based techniques hold substantial promise toward creating digital twins, considering the intrinsic characteristics of railway infrastructure. At the end of this chapter a road map is proposed to aid future research into digitising the rail scene.

Insights of this chapter aided defining the gap identification contribution of this thesis.

\clearpage
\input{./Chapters/stoa/introduction.tex}
\input{./Chapters/stoa/review-method.tex}
\input{./Chapters/stoa/meta_analysis}
\input{./Chapters/stoa/pre-processing}
\input{./Chapters/stoa/modelling-techniques}
\input{./Chapters/stoa/digital-twin}
\input{./Chapters/stoa/discussion.tex}
\input{./Chapters/stoa/conclusion.tex}

\clearpage
\printbibliography[heading=subbibnumbered]